An improved metric for the analysis of swarms

### 2017-11-06 paper.tex
Metric Name Change

%\section{Introduction\label{methods:SwarmStability}}
\emph{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \emph{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively. These rules can also generate movement that does not contribute to the goal of the swarm this movement is a direct result of the agents attempting to move towards an optimum position within the structure. This internal movement can be identified by analysing the changes in the inter-agent interactions. Measuring this unnecessary movement allows the effectivemness of a swarming algorithm to be evaluated. If the internal movement can be reduced then stability of the swarm is improved and more of the swarms energy is focused on the goal. Also reducing the internal movement also provides a more stable platform for the deployment of sensor arrays. Both metric allow this evaluation to be caried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent varies producing varyig distances even though the interactions at the magnitude level indicates the agents are in fact stable and optimum. This paper discusses two metrics and how they differ in their approach to identify these changes. The distance metric uses variations in the inter-agent spaces, as used by Navarro et al.~\cite{NIM:09}. The new metric uses the resultant magnitudes from an agents' \textit{inter-agent vectors} that are induced by an agent interacting with its neighbours which allows a more acurate analysis of a swarm when using non homogenous sensor distances The metric also identifies when a swarm is expanding or contracting.

@@

%% \section{Metric principles\label{section:MagnitudeDynamics}}
%% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
\section{Swarm modelling}\label{SwarmModelling}
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects to implement the swarming characteristic. These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding. Field effects are the ranges around an agent that determine the effect other agents have upon its movement~(Figure~\ref{methods:FieldEffects}). It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$. When an agent ($b'$) moves into the \textit{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the repulsion field of $b$ then $b$ has a tendency to move away from $b'$, i.e. to be repulsed. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has a tendency to move away from the obstacle.

@@

\begin{equation}
\label{eq:Repulsion1}
v_{r}(b) =‎ -
\frac{1}{|rep(b)|}
\left(
\mathlarger{\mathlarger{\sum_{b' \in rep(b)}}}
{\left( 1-\frac{|bb'|}{R_b} \right)}
{bb'}
\right)
\end{equation}‎

@@

If a weighted model is used then equation~\ref{eq:InterAgentMovement1} becomes equation~\ref{eq:InterAgentMovement2}.

@@

\section{Internal movement and the null vector}\label{metric:StabilityNullVector}
When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector. This indicates that two agents are optimally spaced for a given set of conditions. Although the agents are at an optimum position it does not mean the swarm is optimally distributed. If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect. This phenomenon is used in the identification of the emergent behaviour of \textit{area flooding}.

@@

Due to $v(b)$ the independent nature of the agents this situation is very rare. The residual motion that persists in a swarm is the background `noise' or `jitter' that an algorithm creates.

@@

\section{Residual internal movement (Jitter)}\label{metric:Jitter}
Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely. The agent pairs may fluctuate between the 4 states~(Figures~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}). This transitions are \textit{jitter}. The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$ (Equation~\ref{eq:BotStabilityT}) between the agents. Jitter arises as motion to maintain the structure of a swarm. A swarm coordination algorithm that reduces jitter is generally desirable.

@@

\section{Magnitude based metric\label{section:MagnitudeDynamics}}
Magnitude based internal movement (\textit{agent resultant magnitude}) is measured by identifying the balance between the repulsion and cohesion between agents. `Jitter' in the case of the \textit{agent resultant magnitude} metric is measured as the deviation of the magnitude lengths created by the agents. The identification of this deviation produces the clarifying part of the \textit{agent resultant magnitude} metric.
%% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
The \textit{agent resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm. There are two ways of using the cohesion and repulsion in identifying a resultant vector. The two vectors can be added as absolute values to give an overall `size' to the magnitude that is affecting each relationship. Alternatively the resultant magnitude can be the sum of the actual magnitudes. The repulsion vector has a negative magnitude and the cohesion vector has a positive magnitude. In this paper the magnitude analysis will be based on summing the two vectors to determine the result of the inter-agent interaction, although both mechanisms could be used for a comparitive analysis. This paper will refer to the resultant magnitude as the \text{`agent resultant magnitude'} of the relationship. The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude. the \text{`agent resultant magnitude'} can be considered part of the `quality' measure for a swarm's performance.

>>>>

%\section{Introduction\label{methods:SwarmStability}}
\emph{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \emph{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively. These rules can also generate movement that does not contribute to the goal of the swarm this movement is a direct result of the agents attempting to move towards an optimum position within the structure. This internal movement can be identified by analysing the changes in the inter-agent interactions. Measuring this unnecessary movement allows the effectivemness of a swarming algorithm to be evaluated. If the internal movement can be reduced then stability of the swarm is improved and more of the swarms energy is focused on the goal. Also reducing the internal movement also provides a more stable platform for the deployment of sensor arrays. Both metric allow this evaluation to be caried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent varies producing varyig distances even though the interactions at the magnitude level indicates the agents are in fact stable and optimum. This paper discusses two metrics and how they differ in their approach to identify these changes. The \emph{distance metric} which uses variations in the inter-agent distances, as discussed by Navarro et al.~\cite{NIM:09}. The new metric uses the resultant magnitudes of the agents' \textit{inter-agent vectors} that are induced by an agent interacting with its neighbours which allows a more acurate analysis of a swarm when using non homogenous sensor distances The metric also identifies when a swarm is expanding or contracting.

@@

%% \section{Metric principles\label{section:MagnitudeDynamics}}
%% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
\section{Swarm modelling}\label{SwarmModelling}
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects to implement the swarming characteristic. These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding. \emph{Field effects} are the ranges around an agent that determine the effect other agents have upon its movement~(Figure~\ref{methods:FieldEffects}). It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$. When an agent ($b'$) moves into the \textit{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the repulsion field of $b$ then $b$ has a tendency to move away from $b'$, i.e. to be repulsed. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has a tendency to move away from the obstacle.

@@

\begin{equation}
\label{eq:Repulsion1}
v_{r}(b) =‎ 
\frac{1}{|rep(b)|}
\left(
\mathlarger{\mathlarger{\sum_{b' \in rep(b)}}}
{\left( 1-\frac{|bb'|}{R_b} \right)}
{bb'}
\right)
\end{equation}‎

@@

If a \emph{weighted model} is used then equation~\ref{eq:InterAgentMovement1} becomes equation~\ref{eq:InterAgentMovement2}.

@@

\section{Internal movement and the null vector}\label{metric:StabilityNullVector}
When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions. Although the agents are at an optimum position it does not mean the swarm is optimally distributed. If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect. This phenomenon is used in the identification of the emergent behaviour of \textit{area flooding}.

@@

Due to the movement of agents being based only on neighbour interactions this situation is very rare. The residual motion that persists in a swarm is the background `noise' or `jitter' that an algorithm creates.

@@

\section{Residual internal movement (Jitter)}\label{metric:Jitter}
Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely. The agent pairs may fluctuate between the 4 states~(Figures~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \textit{jitter}. The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$ (Equation~\ref{eq:BotStabilityT}) between the agents. Jitter arises as motion to maintain the structure of a swarm. A swarm coordination algorithm that reduces jitter is generally desirable.

@@

\section{Repulsion/Cohesion based metric\label{section:MagnitudeDynamics}}
Repulsion/Cohesion based internal movement (\textit{agent resultant magnitude}) is measured by identifying the balance between the repulsion and cohesion between agents. `Jitter' in the case of the \textit{agent resultant Repulsion/Cohesion} metric is measured as the deviation of the magnitude lengths created by the agents. The identification of this deviation produces the clarifying part of the \textit{agent resultant Repulsion/Cohesion} metric.
%% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
The \textit{agent resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm. The cohesion and repulsion between an agent $b$ and its neighbour $b'$ produce a resultant vector. The vector is calculated by adding the cohesion effect and the repulsion effect. The resultant vector's magnitiude is used as a scalar values to give an overall `size' to the agent relationship. The resultant scalar value is then modified based upon the `size' of the two relationship magnitudes.

==========================

This paper will refer to the scalarised resultant magnitude as the \emph{`resultant cohsion/repulsion metric'} of the relationship. The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude. the \text{`agent resultant magnitude'} can be considered part of the `quality' measure for a swarm's performance.
